CF Registry Analysis

### 2023-05-18 Christine Chan/CF Registry/cf_registry_analysis.qmd
added logistic regression to CFRD registry data

```{r}
# All data
encounter <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
                      na.strings = c("", "N/A")
)
annual <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
                   na.strings = c("", "N/A")
)
encounter <- encounter %>%
  arrange(eDWID, reviewyear, encounternum)
annual = annual %>% arrange(eDWID,ReviewYear)
# Add age
ages <- annual %>% 
  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd)) %>%
  rename(reviewyear = ReviewYear) %>%
  group_by(eDWID,reviewyear) %>%
  summarise(`Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), 
                                                  decreasing = TRUE)[1])),
            Age = Age_YrEnd[1],.groups = "drop")
encounter = left_join(encounter,ages,by = c("eDWID", "reviewyear"))
encounter$`Age Group 2018` = cut(encounter$`Age in 2018`,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
encounter$`Age Group` = cut(encounter$Age,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
# For those with CFRD by OGTT, find first visit with diagnosis
first_cfrd_ogtt <- encounter %>% 
  filter(twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_ogtt = reviewyear)
# For those with CFRD by registry criteria, find first visit with diagnosis
first_cfrd_cffpr <- encounter %>% 
  filter(cfrd_status==3) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_cffpr = reviewyear)
# Both
first_cfrd_both <- encounter %>% 
  filter(cfrd_status==3|twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_both = reviewyear)
# Add to encounter level data
encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age in 2018"))
# Fill down CFRD Yes/No by year
encounter <- encounter %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_ogtt = case_when(
      reviewyear < first_cfrd_ogtt ~ "No",
      reviewyear >= first_cfrd_ogtt ~ "Yes"
    ),
    cfrd_cffpr = case_when(
      reviewyear < first_cfrd_cffpr ~ "No",
      reviewyear >= first_cfrd_cffpr ~ "Yes"
    ),
    cfrd_both = case_when(
      reviewyear < first_cfrd_both ~ "No",
      reviewyear >= first_cfrd_both ~ "Yes"
    )
  ) %>%
  ungroup()
encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
# Ordered factors
encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
encounter$cfrd_cffpr = factor(encounter$cfrd_cffpr,levels = c("No","Yes"),ordered = T)
encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
# Remove those missing age
encounter = encounter %>% filter(!is.na(Age))
```

@@

>>>>

```{r}
# All data
encounter <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
                      na.strings = c("", "N/A")
)
annual <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
                   na.strings = c("", "N/A")
)
encounter <- encounter %>%
  arrange(eDWID, reviewyear, encounternum)
annual = annual %>% arrange(eDWID,ReviewYear)
# Add age
ages <- annual %>% 
  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd)) %>%
  rename(reviewyear = ReviewYear) %>%
  group_by(eDWID,reviewyear) %>%
  summarise(`Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), 
                                                  decreasing = TRUE)[1])),
            Age = Age_YrEnd[1],.groups = "drop")
encounter = left_join(encounter,ages,by = c("eDWID", "reviewyear"))
encounter$`Age Group 2018` = cut(encounter$`Age in 2018`,
                                 c(-Inf,10,19,31,Inf),
                                 labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                                 right=F)
encounter$`Age Group` = cut(encounter$Age,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
# For those with CFRD by OGTT, find first visit with diagnosis
first_cfrd_ogtt <- encounter %>% 
  filter(twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_ogtt = reviewyear)
# For those with CFRD by registry criteria, find first visit with diagnosis
first_cfrd_cffpr <- encounter %>% 
  filter(cfrd_status==3) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_cffpr = reviewyear)
# Both
first_cfrd_both <- encounter %>% 
  filter(cfrd_status==3|twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_both = reviewyear)
# Add to encounter level data
encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age in 2018"))
# Fill down CFRD Yes/No by year
encounter <- encounter %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_ogtt = case_when(
      reviewyear < first_cfrd_ogtt ~ "No",
      reviewyear >= first_cfrd_ogtt ~ "Yes"
    ),
    cfrd_cffpr = case_when(
      reviewyear < first_cfrd_cffpr ~ "No",
      reviewyear >= first_cfrd_cffpr ~ "Yes"
    ),
    cfrd_both = case_when(
      reviewyear < first_cfrd_both ~ "No",
      reviewyear >= first_cfrd_both ~ "Yes"
    )
  ) %>%
  ungroup()
encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
# Ordered factors
encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
encounter$cfrd_cffpr = factor(encounter$cfrd_cffpr,levels = c("No","Yes"),ordered = T)
encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
# Remove those missing age
encounter = encounter %>% filter(!is.na(Age))
```

@@

# Logistic analyses

Any participants with T1D or T2D were excluded from these analyses.

```{r}
# Import and clean data
annual <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
                   na.strings = c("", "N/A")
)
annual = annual %>% filter(cfrd_status_annual != 4,cfrd_status_annual != 5)
annual$cfrd_status_annual = factor(annual$cfrd_status_annual,levels = 1:3,
                                   labels = c("CFRD-","CFRD-","CFRD+"))

demogs = read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_DemogCFDiag_Del1.csv")
demogs = demogs %>% select(eDWID,Gender,MutClass)

annual = left_join(annual,demogs,by = join_by(eDWID))

encounter <- read.csv("/Users/timvigers/Documents/Work/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
                      na.strings = c("", "N/A")
)
# mods = encounter %>% group_by(eDWID,reviewyear) %>%
#   summarise(across(c(Vx770,Vx809comb),~max(.x,na.rm = T)),.groups = "drop")
# Analysis function
log_reg = function(year = 2005){
  # Clean data
  df = annual %>% 
    filter(ReviewYear==year) %>%
    select(eDWID,Age_YrEnd,Gender,MutClass,A_IsOnEnzymes,A_FEV1_pct_predicted,
           A_bmivalue,A_bmipercentile,A_height,A_heightpercentile,
           NumPulmExacerbation,A_feeding2,A_pseudomonasaeruginosa,
           A_fungalyeast1,A_pulmonarycomplications1,A_mycocultureresults1,
           A_corticosteroids1,cfrd_status_annual)
  # Format columns
  df$MutClass = factor(df$MutClass,levels = c("1-3","4-5","Oth","N/A"),
                       labels = c("1-3","4-5","Oth",NA))
  df$NumPulmExacerbation=cut(df$NumPulmExacerbation,labels = c("0","1","2",">=3"),
                             breaks = c(0,1,2,3,Inf),right = F)
  df$A_feeding2 = factor(df$A_feeding2,levels = 0:1,labels = c("No","Yes"))
  df$A_IsOnEnzymes = factor(df$A_IsOnEnzymes,levels = 0:1,labels = c("No","Yes"))
  df$A_FEV1_pct_predicted = cut(df$A_FEV1_pct_predicted,breaks = c(0,40,80,100,Inf),right = F)
  df$A_corticosteroids1 = factor(df$A_corticosteroids1,levels = 0:1,labels = c("No","Yes"))
  df$A_fungalyeast1 = factor(df$A_fungalyeast1,levels = 0:1,labels = c("No","Yes"))
  df$A_pulmonarycomplications1 = factor(df$A_pulmonarycomplications1,levels = 0:1,labels = c("No","Yes"))
  df$A_mycocultureresults1 = factor(df$A_mycocultureresults1,levels = 0:1,labels = c("No","Yes"))
  # Models
  base_mod = as.formula("cfrd_status_annual~Gender+Age_YrEnd+A_IsOnEnzymes+MutClass+
                   A_FEV1_pct_predicted+NumPulmExacerbation+A_feeding2")
  adult_mod = update(base_mod,.~.+A_bmivalue+A_height)
  peds_mod = update(base_mod,.~.+A_bmipercentile+A_heightpercentile)
  # Simple models
  adult_mod1 = glm(adult_mod,data = df,family = "binomial")
  peds_mod1 = glm(peds_mod,data = df,family = "binomial")
  # Add bugs
  # P. aeruginosa
  adult_mod2 = glm(update(adult_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")
  peds_mod2 = glm(update(peds_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")
  # Aspergillus
  adult_mod3 = glm(update(adult_mod,.~.+A_fungalyeast1),data = df,family = "binomial")
  peds_mod3 = glm(update(peds_mod,.~.+A_fungalyeast1),data = df,family = "binomial")
  # ABPA
  adult_mod4 = glm(update(adult_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
  peds_mod4 = glm(update(peds_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
  # Mycobacterium
  adult_mod5 = glm(update(adult_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
  peds_mod5 = glm(update(peds_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")
  # Print results
  cat("\n")
  cat("### Base Models")
  cat("\n")
  cat("#### Adult")
  print(forest_model(adult_mod1))
  cat("\n")
  cat("#### Peds")
  print(forest_model(peds_mod1))
  cat("\n")
  cat("\n")
  cat("### P. aeruginosa")
  cat("\n")
  cat("#### Adult")
  print(forest_model(adult_mod2))
  cat("\n")
  cat("#### Peds")
  print(forest_model(peds_mod2))
  cat("\n")
  cat("\n")
  cat("### Aspergillus")
  cat("\n")
  cat("#### Adult")
  print(forest_model(adult_mod3))
  cat("\n")
  cat("#### Peds")
  print(forest_model(peds_mod3))
  cat("\n")
  cat("\n")
  cat("### ABPA")
  cat("\n")
  cat("#### Adult")
  print(forest_model(adult_mod4))
  cat("\n")
  cat("#### Peds")
  print(forest_model(peds_mod4))
  cat("\n")
  cat("\n")
  cat("### Mycobacterium")
  cat("\n")
  cat("#### Adult")
  print(forest_model(adult_mod5))
  cat("\n")
  cat("#### Peds")
  print(forest_model(peds_mod5))
  cat("\n")
}
```

## 2005

```{r}
log_reg(2005)
```

# Questions

1. What variables do we use for B. Cepacia and liver disease?
2. Do we have CFTR modulator data aside from `Vx770` and `Vx809comb`?
